\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-08 paper/paper 2/dc-2015.tex
2.4

%\emph{Complex constraints} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
Complex constraints of the constraint type \emph{context-specific exclusive or of property groups} (\emph{R-13}) restrict individuals of given classes to have all properties of exactly one of multiple mutually exclusive property groups. Publications, e.g., are either identified by an \emph{ISBN} and a title (for books) or by an \emph{ISSN} and a title (for periodical publications), but it should not be possible to assign both identifiers to a given publication. As \emph{The-Great-Gatsby} is a publication with an \emph{ISBN} and a title without an \emph{ISSN}, \emph{The-Great-Gatsby} is considered as a valid publication. This complex constraint is generically expressible in DL:
%This \emph{complex constraint} is expressible specifically by \emph{ShEx} and generically by DL (and can therefore be mapped to the \emph{RDF-CV} \cite{BoschEckert2015-2}).

@@

%Even though, tools may generate \emph{generic constraints} automatically, this 
The DL statements demonstrate that the complex constraint is composed of many other complex constraints (e.g., \emph{minimum and maximum qualified cardinality restrictions}) and simple constraints (e.g., constraints on sets). As \emph{exact (un)qualified cardinality restrictions (=n)} and \emph{exclusive or (XOR)} are frequently used complex constraints, we propose to simplify them as simple constraints in terms of syntactic sugar. As a consequence, the \emph{context-specific exclusive or of property groups} complex constraint is represented as a generic constraint by means of the RDF-CV more intuitively and concisely (see Table \ref{tab:simplified-complex-constraints}).

@@

Keys, however, are even more general, i.e., a generalization of inverse functional properties \cite{Schneider2009}.
A key can be a datatype, an object property, or a chain of properties.
For these generalization purposes, as there are different sorts of keys, and as keys can lead to undecidability, 
DL is extended with a special construct \emph{keyfor} \cite{Lutz2005}.
When using keyfor (\ms{isbn keyfor Book}), 
the complex constraint can be formulated as a simple constraint which looks like the following in concrete RDF turtle syntax:

@@

Complex constraints of frequently used constraint types which correspond to DL axioms like \emph{transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity} can also be simplified in form of simple constraints. Although, these DL axioms are expressible by basic DL features, they can also be used in terms of syntactic sugar \cite{BoschEckert2015-2}.
%Starfleet officers, e.g., are uniquely identified by their command authorization code (e.g. to activate and cancel auto-destruct sequences).
%It means that the property \emph{commandAuthorizationCode} is inverse functional - mapped to DL and the \emph{RDF-CV} as follows:
%
%\begin{DL}
%$(\ms{funct } commandAuthorizationCode\sp{\overline{\ }})$
%\end{DL}
%
%%\begin{gcotable}
%%property & $\top$ & commandAuthorizationCode$^{-}$ & commandAuthorizationCode$^{-}$ & - & inverse & - \\
%%property & $\top$ & commandAuthorizationCode$^{-}$ & - & - & functional & - \\
%%\end{gcotable}
%
%\begin{table}
  %\scriptsize
  %\sffamily
  %\vspace{0cm}
	%\centering
		%\begin{tabular}{l|l|l|l|l|l|l}
      %\textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      %\hline
%property & $\top$ & commandAuthorizationCode$^{-}$ & commandAuthorizationCode$^{-}$ & - & inverse & - \\
%property & $\top$ & commandAuthorizationCode$^{-}$ & - & - & functional & - \\
		%\end{tabular}
	%\caption{Primary Key Properties as Complex Constraints}
	%\label{tab:primary-key-properties-as-complex-constraints}
%\end{table}
%
%Keys, however, are even more general, i.e., a generalization of inverse functional properties \cite{Schneider2009}.
%A key can be a datatype property, an object property, or a chain of properties.
%For this generalization purposes, as there are different sorts of key, and as keys can lead to undecidability, 
%DL is extended with \emph{key boxes} and a special \emph{keyfor} construct\cite{Lutz2005}.
%This leads to the following DL and \emph{RDF-CV} mappings (only one \emph{simple property constraint}):
%
%\begin{DL}
%commandAuthorizationCode \ms{keyfor} StarfleetOfficer
%\end{DL}

>>>>

%\emph{Complex constraints} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
Complex constraints of the constraint type \emph{context-specific exclusive or of property groups} (\emph{R-13}) restrict individuals of given classes to have all properties of exactly one of multiple mutually exclusive property groups. Publications, e.g., are either identified by an \emph{ISBN} and a title (for books) or by an \emph{ISSN} and a title (for periodical publications), but it should not be possible to assign both identifiers to a given publication. This complex constraint is expressible in \emph{ShEx}:

\begin{ex}
Publication { 
    ( isbn string , title string ) |
    ( issn string , title string ) }
\end{ex}

As \emph{The-Great-Gatsby} is a publication with an \emph{ISBN} and a title without an \emph{ISSN}, \emph{The-Great-Gatsby} is considered as a valid publication. This complex constraint is generically expressible in DL:
%This \emph{complex constraint} is expressible specifically by \emph{ShEx} and generically by DL (and can therefore be mapped to the \emph{RDF-CV} \cite{BoschEckert2015-2}).

@@

%Even though, tools may generate \emph{generic constraints} automatically, this 
The DL statements demonstrate that the complex constraint is composed of many other complex constraints (\emph{minimum (R-75) and maximum qualified cardinality restrictions (R-76)}) and simple constraints (\emph{intersection (R-15/16)}, \emph{disjunction (R-17/18)}, and \emph{negation (R-19/20)}). Constraints of almost 14\% of the constraint types are complex constraints which can be simplified and therefore formulated as simple constraints when using them in terms of syntactic sugar. As \emph{exact (un)qualified cardinality restrictions (=n)} and \emph{exclusive or (XOR) of property groups} are frequently used complex constraints, we propose to simplify them in form of simple constraints. As a consequence, the \emph{context-specific exclusive or of property groups (R-13)} complex constraint is represented as a generic constraint by means of the RDF-CV more intuitively and concisely (see Table \ref{tab:simplified-complex-constraints}).

@@

Keys, however, are even more general, i.e., a generalization of inverse functional properties \cite{Schneider2009}.
A key can be a datatype, an object property, or a chain of properties.
For these generalization purposes, as there are different sorts of keys, and as keys can lead to undecidability, 
DL is extended with a special construct \emph{keyfor} \cite{Lutz2005}.
When using keyfor (\ms{isbn keyfor Book}), 
the complex constraint can be simplified and thus formulated as a simple constraint which looks like the following in concrete RDF turtle syntax:

@@

Complex constraints of frequently used constraint types which correspond to DL axioms like \emph{transitivity}, \emph{symmetry}, \emph{asymmetry}, \emph{reflexivity} and \emph{irreflexivity} can also be simplified in form of simple constraints. Although, these DL axioms are expressible by basic DL features, they can also be used in terms of syntactic sugar \cite{BoschEckert2015-2}.

Constraints of the \emph{irreflexive object properties (R-60)} constraint type ensure that no individual is connected by a given object property to itself \cite{Kroetzsch2012}. With the irreflexive object property constraint {\small\ms{$\top$ $\sqsubseteq$ $\neg$ $\exists authorOf.Self$}}, e.g., one can state that individuals cannot be authors of themselves. When represented using the RDF-CV, the complex constraint aggregates three simple constraints - one property and two class constraints (see Table \ref{tab:irreflexive-object-properties-as-complex-constraints}).

%\begin{gcotable}
%property & $\exists$ marriedTo . Self & marriedTo & - & Self & $\exists$ & - \\
%class & $\neg$ $\exists$ marriedTo . Self & - & - & $\exists$ marriedTo . Self & $\neg$ & - \\
%class & $\top$ & - & - & $\top$, $\neg$ $\exists$ marriedTo . Self & $\sqsubseteq$ & - \\
%\end{gcotable}

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Irreflexive Object Properties as Complex Constraints}
	\label{tab:irreflexive-object-properties-as-complex-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
property & $\exists$ authorOf.Self & authorOf & - & Self & existential quantification & - \\
class & $\neg$ $\exists$ authorOf.Self & - & - & $\exists$ authorOf.Self & negation & - \\
class & $\top$ & - & - & $\top$, $\neg$ $\exists$ authorOf.Self & sub-class & - \\
		\end{tabular}
\end{table}

When using the irreflexive object property constraint in terms of syntactic sugar, 
the complex constraint can be expressed more concisely in form of a simple property constraint with exactly the same semantics (see Table \ref{tab:irreflexive-object-properties-as-simple-constraints}):

%\begin{gcotable}
%property & $\top$ & marriedTo & - & - & irreflexive & - \\
%\end{gcotable}

\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Irreflexive Object Properties as Simple Constraints}
	\label{tab:irreflexive-object-properties-as-simple-constraints}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
property & $\top$ & authorOf & - & - & irreflexive property & - \\
		\end{tabular}
\end{table}

%Starfleet officers, e.g., are uniquely identified by their command authorization code (e.g. to activate and cancel auto-destruct sequences).
%It means that the property \emph{commandAuthorizationCode} is inverse functional - mapped to DL and the \emph{RDF-CV} as follows:
%
%\begin{DL}
%$(\ms{funct } commandAuthorizationCode\sp{\overline{\ }})$
%\end{DL}
%
%%\begin{gcotable}
%%property & $\top$ & commandAuthorizationCode$^{-}$ & commandAuthorizationCode$^{-}$ & - & inverse & - \\
%%property & $\top$ & commandAuthorizationCode$^{-}$ & - & - & functional & - \\
%%\end{gcotable}
%
%\begin{table}
  %\scriptsize
  %\sffamily
  %\vspace{0cm}
	%\centering
		%\begin{tabular}{l|l|l|l|l|l|l}
      %\textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      %\hline
%property & $\top$ & commandAuthorizationCode$^{-}$ & commandAuthorizationCode$^{-}$ & - & inverse & - \\
%property & $\top$ & commandAuthorizationCode$^{-}$ & - & - & functional & - \\
		%\end{tabular}
	%\caption{Primary Key Properties as Complex Constraints}
	%\label{tab:primary-key-properties-as-complex-constraints}
%\end{table}
%
%Keys, however, are even more general, i.e., a generalization of inverse functional properties \cite{Schneider2009}.
%A key can be a datatype property, an object property, or a chain of properties.
%For this generalization purposes, as there are different sorts of key, and as keys can lead to undecidability, 
%DL is extended with \emph{key boxes} and a special \emph{keyfor} construct\cite{Lutz2005}.
%This leads to the following DL and \emph{RDF-CV} mappings (only one \emph{simple property constraint}):
%
%\begin{DL}
%commandAuthorizationCode \ms{keyfor} StarfleetOfficer
%\end{DL}
